Notes for the week of 4/27/19 - 5/3/19

### 2019-05-13 neurips/local_spectral_density_clustering.tex
Added weekly notes detailing TV to infinity mixing time for lazy random walk

%%%%%%%% Local Spectral Clustering of Density Upper Level Sets %%%%%%%%%%%%%%%%%

\documentclass{article}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{float}
\usepackage[export]{adjustbox}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{xr-hyper}
\usepackage{hyperref}
\usepackage[reqno]{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage[parfill]{parskip}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{bm}
\usepackage{fullpage}

\usepackage[square,sort,comma,numbers]{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

% Left and right equations
\makeatletter
\newcommand{\leqnomode}{\tagsleft@true\let\veqno\@@leqno}
\newcommand{\reqnomode}{\tagsleft@false\let\veqno\@@eqno}
\makeatother

@@

@@

\title{Local Spectral Clustering of Density Upper Level Sets}
\author{Alden Green, Sivaraman Balakrishnan, and Ryan Tibshirani}
\date{\today}
\maketitle

\begin{abstract}
\vskip 0.1 in % RESOLVED
Spectral clustering methods are a family of popular nonparametric clustering
tools.  Recent works have proposed and analyzed \emph{local} spectral methods,
which extract clusters using locally-biased random walks around a user-specified
seed node.  Several authors have shown that local methods, such as personalized
PageRank (PPR), have worst-case guarantees for certain graph-based measures of
cluster quality.  In contrast to existing works, we analyze PPR in a traditional
statistical learning setup, where we obtain samples from an unknown
distribution, and aim to identify connected regions of high-density (density
clusters).  We introduce two natural criteria for cluster quality, and derive
bounds for these criteria when evaluated on empirical analogues of density
clusters. Moreover, we prove that PPR, run on a neighborhood graph, extracts
sufficiently salient density clusters.
\end{abstract}

@@

We will outline the main techniques used to prove Theorem \ref{thm: mixing_time_upper_bound}, but first, we show that a straightforward and perhaps immediately obvious route to upper bounding the mixing time is insufficient for our purposes. Let $\widetilde{\Phi}_{n,r}$ be the \emph{conductance} of the induced subgraph $\widetilde{G}_{n,r} = G_{n,r}\bigl[\Csig[\Xbf]\bigr]$, given by
\begin{equation*}
\widetilde{\Phi}_{n,r} = \min_{S \subseteq \Csig[\Xbf]} \Phi(S; \widetilde{G}_{n,r}),
\end{equation*}
and $\widetilde{\mu}_2$ be the 2nd largest eigenvalue of $\Wbf_{\Csig[\Xbf]}$, the random walk matrix over $\widetilde{G}_{n,r}$. The famous Cheeger's inequality
\begin{equation*}
\frac{\widetilde{\Phi}_{n,r}^2}{2} \leq 1 - \widetilde{\mu}_2 \leq \widetilde{\Phi}_{n,r}
\end{equation*}
applied with well known theory on $\infty$-mixing time yields
\begin{equation*}
\abs{\widetilde{\pi}_{n,r}(u) - e_v \Wbf_{\Csig[\Xbf]}^te_u} \leq \left(1 - \frac{\widetilde{\Phi}_{n,r}^2}{2}\right)^t \sqrt{\frac{\deg(u; \widetilde{G}_{n,r})}{\min_{v \in \Csig[\Xbf]} \deg(v; \widetilde{G}_{n,r})}} \tag{$u,v \in \Csig[\Xbf]$}
\end{equation*}
where $\widetilde{\pi}_{n,r}$ is the stationary distribution over $\widetilde{G}_{n,r}$ Unfortunately, even if $\widetilde{G}_{n,r}$ is an approximately degree-regular graph, (meaning we can ignore the term under the square root), this bound only yields \footnote{For sequences of numbers $\set{a_n}_{n \in \mathbb{N}}, \set{b_n}_{n \in \mathbb{N}}$, we say $a_n = \mathcal{O}(b_n)$ when there exists $c$ sufficiently large such that $a_n \leq c b_n$ for all $n$.}
\begin{equation*}
\tau_{n,r}(\Csig[\Xbf]) = \mathcal{O} \left( \frac{\log\left(\frac{1}{\widetilde{\pi}_{\min}}\right)}{\widetilde{\Phi}_{n,r}^2} \right).
\end{equation*}
where $\widetilde{\pi}_{\min} = \min_{u \in \Csig[\Xbf]} \widetilde{\pi}_{n,r}(u)$. As $\widetilde{\pi}_{\min} = \mathcal{O}(1/n)$, the upper bound on the right hand side will be of order $\log(n)$, and the bound becomes trivial as $n \to \infty$.

>>>>

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2019

% ready for submission
% \usepackage{neurips_2019}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2019}

% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2019}

\usepackage{neurips_2019}

% to avoid loading the natbib package, add option nonatbib:
%     \usepackage[nonatbib]{neurips_2019}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{float}
\usepackage[export]{adjustbox}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{xr-hyper}
\usepackage{hyperref}
\usepackage[reqno]{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage[parfill]{parskip}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{bm}

@@

\title{Local Spectral Clustering of Density Upper Level Sets}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{%
  David S.~Hippocampus\thanks{Use footnote for providing further information
    about author (webpage, alternative address)---\emph{not} for acknowledging
    funding agencies.} \\
  Department of Computer Science\\
  Cranberry-Lemon University\\
  Pittsburgh, PA 15213 \\
  \texttt{hippo@cs.cranberry-lemon.edu} \\
  % examples of more authors
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \AND
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

@@

\maketitle

\begin{abstract}
\vskip 0.1 in
  Spectral clustering methods are a family of popular nonparametric clustering
  tools.  Recent works have proposed and analyzed \emph{local} spectral methods,
  which extract clusters using locally-biased random walks around a user-specified
  seed node.  Several authors have shown that local methods, such as personalized
  PageRank (PPR), have worst-case guarantees for certain graph-based measures of
  cluster quality.  In contrast to existing works, we analyze PPR in a traditional
  statistical learning setup, where we obtain samples from an unknown
  distribution, and aim to identify connected regions of high-density (density
  clusters).  We introduce two natural criteria for cluster quality, and derive
  bounds for these criteria when evaluated on empirical analogues of density
  clusters. Moreover, we prove that PPR, run on a neighborhood graph, extracts
  sufficiently salient density clusters. Finally, we provide empirical support of our theory.
\end{abstract}

@@

We will outline the main techniques used to prove Theorem \ref{thm: mixing_time_upper_bound}, but first, we show that a straightforward and perhaps immediately obvious route to upper bounding the mixing time is insufficient for our purposes. Let $\widetilde{\Phi}_{n,r}$ be the \emph{conductance} of the induced subgraph $\widetilde{G}_{n,r} = G_{n,r}\bigl[\Csig[\Xbf]\bigr]$, given by
\begin{equation*}
\widetilde{\Phi}_{n,r} = \min_{S \subseteq \Csig[\Xbf]} \Phi(S; \widetilde{G}_{n,r}),
\end{equation*}
and $\widetilde{\mu}_2$ be the 2nd largest eigenvalue of $\Wbf_{\Csig[\Xbf]}$, the random walk matrix over $\widetilde{G}_{n,r}$. The famous Cheeger's inequality
\begin{equation*}
\frac{\widetilde{\Phi}_{n,r}^2}{2} \leq 1 - \widetilde{\mu}_2 \leq \widetilde{\Phi}_{n,r}
\end{equation*}
applied with well known theory on $\infty$-mixing time yields
\begin{equation*}
\abs{\widetilde{\pi}_{n,r}(u) - e_v \Wbf_{\Csig[\Xbf]}^te_u} \leq \left(1 - \frac{\widetilde{\Phi}_{n,r}^2}{2}\right)^t \sqrt{\frac{\deg(u; \widetilde{G}_{n,r})}{\min_{v \in \Csig[\Xbf]} \deg(v; \widetilde{G}_{n,r})}} \tag{$u,v \in \Csig[\Xbf]$}
\end{equation*}
where $\widetilde{\pi}_{n,r}$ is the stationary distribution over $\widetilde{G}_{n,r}$ Unfortunately, even if $\widetilde{G}_{n,r}$ is an approximately degree-regular graph, (meaning we can ignore the term under the square root), this bound only yields \footnote{For sequences of numbers $\set{a_n}_{n \in \mathbb{N}}, \set{b_n}_{n \in \mathbb{N}}$, we say $a_n = \mathcal{O}(b_n)$ when there exists $c$ sufficiently large such that $a_n \leq c b_n$ for all $n$.}
\begin{equation*}
\tau_{n,r}(\Csig[\Xbf]) = \mathcal{O} \left( \frac{\log\left(\frac{1}{\widetilde{\pi}_{\min}}\right)}{\widetilde{\Phi}_{n,r}^2} \right).
\end{equation*}
where $\widetilde{\pi}_{\min} = \min_{u \in \Csig[\Xbf]} \widetilde{\pi}_{n,r}(u)$. As $\widetilde{\pi}_{\min} = \mathcal{O}(1/n)$, the upper bound on the right hand side will be of order at least $\log(n)$, and the bound becomes trivial as $n \to \infty$.

### 2019-05-13 notes/weekly_notes/4.27.19.NotesForWeek.tex
Added weekly notes detailing TV to infinity mixing time for lazy random walk

	Observe that since $\Aset$ is closed and bounded, it is compact. As $B(x,\sigma)$ is compact, and the direct sum of two compact sets is also compact, $\Asig$ is compact. Moreover,
	\begin{equation*}
	\Asig \subset \bigcup_{x \in \Aset} B^{\circ}(x,\sigma + \epsilon)
	\end{equation*}
	so by compactness there exists $x_1, \ldots,x_N \in \Aset$ such that
	\begin{equation*}
	\Asig \subset \bigcup_{i = 1}^{N} B^{\circ}(x_i,\sigma + \epsilon).
	\end{equation*}

@@

>>>>

	Observe that since $\Aset$ is closed and bounded, it is compact. As $B(x,\sigma)$ is compact, and the direct sum of two compact sets is itself compact, $\Asig$ is compact. Moreover,
	\begin{equation*}
	\Asig \subset \bigcup_{x \in \Aset} B^{\circ}(x,\sigma + \epsilon)
	\end{equation*}
	so by compactness there exists $x_1, \ldots,x_N \in \Aset$ such that
	\begin{equation*}
	\Asig \subset \bigcup_{i = 1}^{N} B^{\circ}(x_i,\sigma + \epsilon).
	\end{equation*}

@@

\section{Old version of the same result.}

For two sets $\mathcal{X},\mathcal{Y}$ and scalar $\alpha$, we write 
\begin{equation*}
\mathcal{X} + \mathcal{Y} := \set{z: z = x + y, x \in \mathcal{X}, y \in \mathcal{Y}},~ \mathcal{X} - \mathcal{Y} := \set{z: z = x - y, x \in \mathcal{X}, y \in \mathcal{Y}}
\end{equation*}
and $\alpha \mathcal{X} = \set{z: z = \alpha x, x \in \mathcal{X}}$.

Let $B = B(0,1)$ be the $d$-dimensional unit ball centered at the origin.

\begin{lemma}
	\label{lem: expansion_sets}
	Let $\mathcal{K} \subseteq \Reals^d$ be a convex set, and $g: \Reals^d \to \Reals^d$ be a measure preserving Lipschitz mapping, meaning there exists $ L > 0$ such that
	\begin{equation*}
	\frac{1}{L} \norm{x - y} \leq \norm{g(x) - g(y)} \leq L \norm{x - y} \tag{for all $x,y \in \Reals^d$}
	\end{equation*}
	and $\norm{\nabla g (x)} = 1$ for all $x \in \Reals^d$. Letting $\mathcal{A} = g(\mathcal{K})$, the following statement holds: if for $\sigma > 0$, there exists some $x \in \Rd$ such that 
	\begin{equation}
	\label{eqn: expansion_sets}
	B(x,\sigma) \subseteq \mathcal{A} 
	\end{equation}
	then for any $\delta > 0$
	\begin{equation*}
	\nu(\mathcal{A} + \delta B) \leq \left(1 + \frac{L^2 \delta}{\sigma}\right)^d \nu(\mathcal{A})
	\end{equation*}
\end{lemma}
\begin{proof}
	We begin by showing the statement in the convex case, where $g = \mathrm{Id}$ is the identity mapping, before proceeding to show it in full generality. Without loss of generality assume $\sigma B \subseteq \mathcal{A}$, otherwise center by $\mathcal{A} - x$ for some $x$ such that $B(x,\sigma) \subseteq \mathcal{A}$, and as Lebesgue measure is invariant under translation any bound on the volume of $\mathcal{A} - x$ holds for $\mathcal{A}$ as well.

	\textbf{Convex case, $g = \Id$.}

	Note that 
	\begin{equation}
	\label{eqn: expansion_sets_1}
	\mathcal{A} + \delta B \subseteq \mathcal{A} + \frac{\delta}{\sigma}\mathcal{A}
	\end{equation}
	as $y \in \delta B$ implies $\frac{\sigma}{\delta}y \in \sigma B \subseteq \mathcal{A}$. For $z \in \mathcal{A} + \frac{\delta}{\sigma}\mathcal{A}$, write $z := a + \frac{\delta}{\sigma} a'$ for some $a,a' \in \mathcal{A}$. Then
	\begin{equation}
	\label{eqn: expansion_sets_2}
	\frac{1}{1 + \delta/\sigma}z = \frac{1}{1 + \delta/\sigma}a + \frac{\delta/\sigma}{1 + \delta/\sigma} a' \in \mathcal{A}
	\end{equation}
	where the last statement follows by the convexity of $\mathcal{A}$. Together, \eqref{eqn: expansion_sets_1} and \eqref{eqn: expansion_sets_2} imply $\mathcal{A} + \delta B \subseteq (1 + \frac{\delta}{\sigma})\mathcal{A}$, and therefore
	\begin{equation*}
	\nu(\mathcal{A} + \delta B) \leq \nu\Biggl(\left(1 + \frac{\delta}{\sigma}\right)\mathcal{A}\Biggr) = \left(1 + \frac{\delta}{\sigma}\right)^d \nu(\mathcal{A})
	\end{equation*}
	where the last statement follows by the change of variables formula.

	\textbf{Non-convex case, $g$ general.} For any $z \in \mathcal{A} + \delta B$, there exists $x \in \mathcal{A}$ such that $\norm{z - x} \leq \delta$. Therefore
	\begin{equation*}
	\norm{g^{-1}(z) - g^{-1}(x)} \leq L \delta
	\end{equation*}
	and since $g^{-1}(x) \in \mathcal{K}$, $g^{-1}(z) \in \mathcal{K} + L \delta B$. As a result, $\mathcal{A} + \delta B \subseteq g(\mathcal{K} + L \delta B)$; since $g$ is measure-preserving, we have
	\begin{equation*}
	\nu(\mathcal{A} + \delta B) \leq \nu(\mathcal{K} + L \delta B)
	\end{equation*}

	Additionally, note that the hypothesis $\sigma B \subseteq \mathcal{A}$ implies that $g^{-1}(\sigma B) \subseteq \mathcal{K}$, and if $\norm{z} \leq \frac{\sigma}{L}$, then $g(z) \in \sigma(B)$ and $z \in g^{-1}(\sigma B)$. Therefore, letting $\sigma' = \frac{\sigma}{L}$, we have
	\begin{equation*}
	\sigma' B \subseteq \mathcal{K}
	\end{equation*}
	and the general result follows from applying the convex result to $\mathcal{K} + \delta'B$ for $\delta' = \delta L$. 
\end{proof}
